---
title: "Project 3 - Survey results for Language, Database, and Platform Preference"
author: "Thomas Hill"
date: "3/19/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Languages

```{r import-lang-csv}
survey_lang_raw <- read.csv("C:/Users/Thomas/Documents/GitHub/DATA607_Project_3/LanguageSkills.csv")

head(survey_lang_raw)
names(survey_lang_raw)
```
The languages dataframe contains 25 columns to rate a survey respondent's languages. I assume that the majority of people have only ranked the 3-5 languages. I'll check my assumption by counting the amount of 'NA' entries, then eliminate the unnecessary columns.


```{r lang-remove-empty-columns}
sapply(survey_lang_raw, function(x) sum(is.na(x)))
survey_lang_remove_na <- subset(survey_lang_raw, is.na(survey_lang_raw$Language1) == FALSE)
round(1-sapply(survey_lang_remove_na, function(x) sum(is.na(x))) /sum(is.na(survey_lang_remove_na$Language25)),3)
```
After removing all 'NA''s in the first language column, which coincide with non-response to this question, it appears that the last 6 columns are not contributing appreciably. The majority of responndents listed 5 languages they commonly used.

Next, I'm going to gather the Languages to see the most common overall responses were. This will sum all responses regardless of rank. I'm also going to use the facet wrap function in ggplot to 

```{r lang-plot-all}
library(tidyr)
library(dplyr)
library(ggplot2)

lang_tall <- survey_lang_remove_na %>%
  select(Language1:Language19) %>%
  gather(LangRank, Language, na.rm = TRUE)
lang_wide <- lang_tall %>%
  select(Language) %>%
  group_by(Language) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))

head(lang_wide)
nrow(lang_wide)
lang_wide[[2]][28]/lang_wide[[2]][1]

plot_lang_freq <- head(lang_wide, 10) %>%
  ggplot(aes(x = reorder(Language, Count), y = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Count), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 45))

plot_lang_freq
```

```{r lang-facet-wrap}

lang_tall_trim <- survey_lang_remove_na %>%
  select(Language1:Language10) %>%
  gather(LangRank, Language, na.rm = TRUE) 

tail(lang_tall_trim)

plot_lang_wrap_trim <- lang_tall_trim %>%
  ggplot(aes(x= Language)) +
  geom_bar() +
  facet_wrap(~LangRank)
plot_lang_wrap_trim
```
The most insteresting finding from the facet wrap is that there is one language early in the alphabet overrepresented at Language1, and two languagues in Language2 and 3 that are especially prominent. To represent this data for the presentation, I'll create plots for these three languages sepecifically.


```{r plot-first-language-preference}

lang_1_tall <- survey_lang_remove_na %>%
  select(Language1) %>%
  gather(LangRank, Language, na.rm = TRUE)
lang_1_wide <- lang_1_tall %>%
  select(Language) %>%
  group_by(Language) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))

plot_lang_1_freq <- head(lang_1_wide) %>%
  ggplot(aes(x = reorder(Language, Count), y = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Count), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 45))
plot_lang_1_freq
```

```{r plot-second-language-preference}

lang_2_tall <- survey_lang_remove_na %>%
  select(Language2) %>%
  gather(LangRank, Language, na.rm = TRUE)
lang_2_wide <- lang_2_tall %>%
  select(Language) %>%
  group_by(Language) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))

plot_lang_2_freq <- head(lang_2_wide) %>%
  ggplot(aes(x = reorder(Language, Count), y = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Count), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 45))
plot_lang_2_freq
```

```{r plot-third-language-preference}

lang_3_tall <- survey_lang_remove_na %>%
  select(Language3) %>%
  gather(LangRank, Language, na.rm = TRUE)
lang_3_wide <- lang_3_tall %>%
  select(Language) %>%
  group_by(Language) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))

plot_lang_3_freq <- head(lang_3_wide) %>%
  ggplot(aes(x = reorder(Language, Count), y = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Count), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 45))
plot_lang_3_freq
```

Looking at the top three languages, Powershell, HTML/CSS, and Javascript are the most common in the 3 language preferences. This is an unintuitive finding as overall, SQL and Python are the most common languages.  AS a final consideration, I'll take a second look the most common languages by looking only at the top language preferences and see if there are some languages with entries spread out between the top five.

```{r plot-lang-top-five}

lang_tall_5 <- survey_lang_remove_na %>%
  select(Language1:Language5) %>%
  gather(LangRank, Language, na.rm = TRUE)
lang_wide_5 <- lang_tall_5 %>%
  select(Language) %>%
  group_by(Language) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))

plot_lang_freq_top_5 <- head(lang_wide_5) %>%
  ggplot(aes(x = reorder(Language, Count), y = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Count), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 45))

plot_lang_freq_top_5
```

So while PowerShell is developers' favorite first pick, it's actually only the fourth most common pick for the top five programming languages.  HTML/CSS is the top preference in the 5 first languages, followed by Javascript, and then Python edges out Powershell as the third most common.  


# Databases

```{r import-db-csv}
survey_db_raw <- read.csv("C:/Users/Thomas/Documents/GitHub/DATA607_Project_3/DatabaseSkills.csv")

head(survey_db_raw)
names(survey_db_raw)
```
Databases preference contains another 13 columns, I will check the same assumption about superfluous columns.

```{r db-remove-empty-columns}
sapply(survey_db_raw, function(x) sum(is.na(x)))
survey_db_remove_na <- subset(survey_db_raw, is.na(survey_db_raw$Database1) == FALSE)

#sapply(survey_db_remove_na, function(x) sum(is.na(x))) - sapply(survey_db_raw, function(x) sum(is.na(x)))

round(1-sapply(survey_db_remove_na, function(x) sum(is.na(x))) /sum(is.na(survey_db_remove_na$Database13)),3)
```

It appears that the last row does not contain any additional information, and the last 5 contain negligible amounts.

```{r db-plot-all}
library(tidyr)
library(dplyr)
library(ggplot2)

db_tall <- survey_db_remove_na %>%
  select(Database1:Database9) %>%
  gather(DbRank, Database, na.rm = TRUE)
db_wide <- db_tall %>%
  select(Database) %>%
  group_by(Database) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))

head(db_wide)
nrow(db_wide)
db_wide[[2]][14]/db_wide[[2]][1]

plot_db_freq <- head(db_wide, 10) %>%
  ggplot(aes(x = reorder(Database, Count), y = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Count), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 45))

plot_db_freq
```

```{r db-facet-wrap}

db_tall_trim <- survey_db_remove_na %>%
  select(Database1:Database7) %>%
  gather(DbRank, Database, na.rm = TRUE) 

tail(db_tall_trim)
  
plot_db_wrap_trim <- db_tall_trim %>%
  ggplot(aes(x= Database)) +
  geom_bar() +
  facet_wrap(~DbRank)
plot_db_wrap_trim
```

Again, there appear to be three Databases clearly favored in the top three. I'll apply the same exploration and plots to Database as I did in the Language example.

```{r plot-first-db-preference}

db_1_tall <- survey_db_remove_na %>%
  select(Database1) %>%
  gather(DbRank, Database, na.rm = TRUE)
db_1_wide <- db_1_tall %>%
  select(Database) %>%
  group_by(Database) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))

plot_db_1_freq <- head(db_1_wide) %>%
  ggplot(aes(x = reorder(Database, Count), y = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Count), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 45))
plot_db_1_freq
```

```{r plot-second-db-preference}

db_2_tall <- survey_db_remove_na %>%
  select(Database2) %>%
  gather(DbRank, Database, na.rm = TRUE)
db_2_wide <- db_2_tall %>%
  select(Database) %>%
  group_by(Database) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))

plot_db_2_freq <- head(db_2_wide) %>%
  ggplot(aes(x = reorder(Database, Count), y = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Count), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 45))
plot_db_2_freq
```

```{r plot-third-db-preference}

db_3_tall <- survey_db_remove_na %>%
  select(Database3) %>%
  gather(DbRank, Database, na.rm = TRUE)
db_3_wide <- db_3_tall %>%
  select(Database) %>%
  group_by(Database) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))

plot_db_3_freq <- head(db_3_wide) %>%
  ggplot(aes(x = reorder(Database, Count), y = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Count), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 45))
plot_db_3_freq
```


```{r plot-db-top-five}

db_tall_5 <- survey_db_remove_na %>%
  select(Database1:Database5) %>%
  gather(DbRank, Database, na.rm = TRUE)
db_wide_5 <- db_tall_5 %>%
  select(Database) %>%
  group_by(Database) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))

plot_db_freq_top_5 <- head(db_wide_5) %>%
  ggplot(aes(x = reorder(Database, Count), y = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Count), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 45))

plot_db_freq_top_5
```

The number one preferred database for developers is Microsoft SQL Server, although taken as a whole MySQL and PostgreSQL are SQLite and MongoDB, and generaly there appears to be less diversity than programming languages. Only 14 databases are represented as options, as opposed to 28 Languages. The least common database is ~4% of the most commmon response, as compared to 1.9% which is the statistic for Languages.

#Platforms

```{r import-pf-csv}

survey_pf_raw <- read.csv("C:/Users/Thomas/Documents/GitHub/DATA607_Project_3/PlatformSkills.csv")
head(survey_pf_raw)
names(survey_pf_raw)

```

```{r pf-remove-empty-columns}

sapply(survey_pf_raw, function(x) sum(is.na(x)))

survey_pf_remove_na <- subset(survey_pf_raw, is.na(survey_pf_raw$Platform1) == FALSE)

round(1-sapply(survey_pf_remove_na, function(x) sum(is.na(x))) /sum(is.na(survey_pf_remove_na$Platform16)),3)

#sapply(survey_pf_remove_na, function(x) sum(is.na(x))) - sapply(survey_pf_raw, function(x) sum(is.na(x)))
```
Out of 16 possible platforms, most respondents picked their top three.  Platforms 12-16 represent a negligible amount of information to be gained.

```{r pf-plot-all}
library(tidyr)
library(dplyr)
library(ggplot2)

pf_tall <- survey_pf_remove_na %>%
  select(Platform1:Platform11) %>%
  gather(PfRank, Platform, na.rm = TRUE)
pf_wide <- pf_tall %>%
  select(Platform) %>%
  group_by(Platform) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))

head(pf_wide)
nrow(pf_wide)
pf_wide[[2]][17]/pf_wide[[2]][1]

plot_pf_freq <- head(pf_wide, 10) %>%
  ggplot(aes(x = reorder(Platform, Count), y = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Count), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 45))

plot_pf_freq
```

The top three database responses are Linux and Windows, with a dropoff after these two, then Docker, AWS, and MacOS.


```{r pf-facet-wrap}

pf_tall_trim <- survey_pf_remove_na %>%
  select(Platform1:Platform8) %>%
  gather(PfRank, Platform, na.rm = TRUE) 

tail(pf_tall_trim)
  
plot_pf_wrap_trim <- pf_tall_trim %>%
  ggplot(aes(x= Platform)) +
  geom_bar() +
  facet_wrap(~PfRank)
plot_pf_wrap_trim
```

The platform text wrap reveals that for the top three responses, there exist a handful of platforms that are preferred, with a significant dropoff after the third or fourth options.

```{r plot-first-pf-preference}

pf_1_tall <- survey_pf_remove_na %>%
  select(Platform1) %>%
  gather(PfRank, Platform, na.rm = TRUE)
pf_1_wide <- pf_1_tall %>%
  select(Platform) %>%
  group_by(Platform) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))

plot_pf_1_freq <- head(pf_1_wide) %>%
  ggplot(aes(x = reorder(Platform, Count), y = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Count), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 45))
plot_pf_1_freq
```

```{r plot-second-pf-preference}

pf_2_tall <- survey_pf_remove_na %>%
  select(Platform2) %>%
  gather(PfRank, Platform, na.rm = TRUE)
pf_2_wide <- pf_2_tall %>%
  select(Platform) %>%
  group_by(Platform) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))

plot_pf_2_freq <- head(pf_2_wide) %>%
  ggplot(aes(x = reorder(Platform, Count), y = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Count), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 45))
plot_pf_2_freq
```

```{r plot-third-pf-preference}

pf_3_tall <- survey_pf_remove_na %>%
  select(Platform3) %>%
  gather(PfRank, Platform, na.rm = TRUE)
pf_3_wide <- pf_3_tall %>%
  select(Platform) %>%
  group_by(Platform) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))

plot_pf_3_freq <- head(pf_3_wide) %>%
  ggplot(aes(x = reorder(Platform, Count), y = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Count), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 45))
plot_pf_3_freq
```

```{r plot-pf-top-five}

pf_tall_5 <- survey_pf_remove_na %>%
  select(Platform1:Platform5) %>%
  gather(PfRank, Platform, na.rm = TRUE)
pf_wide_5 <- pf_tall_5 %>%
  select(Platform) %>%
  group_by(Platform) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))

plot_pf_freq_top_5 <- head(pf_wide_5) %>%
  ggplot(aes(x = reorder(Platform, Count), y = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Count), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 45))

plot_pf_freq_top_5
```

Looking at the Platform reponses, it's a robust finding that Windows and Linux are the most popular, as they are represented in all of our plots, especially the more aggregated plots.  However, it's worth mentioning that Android and AWS are the top two first pick for many developers, and Docker is a preferred second pick. Again, there's more diversity in the responses for platforms as even the most uncommon answer is ~5% compared to the most common option.  