---
title: "Project 3 - Data Science Skills"
author: "Ken Popkin, Thomas Hill, Patrick Maloney"
date: "3/21/2020"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```


```{r}
library(RMySQL)
library(dplyr)
library(ggplot2)
```


### Read from AWS MySQL database
```{r}
db = dbConnect(MySQL(), user='root', password='masterfightingduckhorse', dbname='project3', host='mydbinstance.c1u7dqmta5za.us-east-1.rds.amazonaws.com', port=3306)
```

```{r}
rs = dbSendQuery(db, "select * from demographics")
dfdemographics = fetch(rs, n=-1)
```

```{r}
dfdemographics <- select (dfdemographics,-c(dummy))
dim(dfdemographics)
```
 
We'll now take a look at job satisfaction for countries with a medium (100 - 999 respondents) and high (> 1000 respondents)
```{r}
dfcountry <- select(dfdemographics, country)

response <- dfcountry %>%
  group_by(country) %>%
  summarise(length(country)) %>%
  rename(respondents = 2)
  
medium_response <- subset(response, respondents > 99 & respondents < 501)
high_response <- subset(response, respondents > 999)

medium_response_vector <- pull(medium_response,country)
high_response_vector <- pull(high_response,country)

cat('Medium response countries are:', medium_response_vector, '\n\n')
cat('High response countries are:', high_response_vector)
```

```{r}
#Dataframe of medium response countries job satisfaction ratings
medium_response2 <- filter(dfdemographics, country %in% medium_response_vector)
dim(medium_response2)

#Dataframe of high response countries job satisfaction ratings
high_response2 <- filter(dfdemographics, country %in% high_response_vector)
dim(high_response2)
```

Remove entries where job satisfaction was not answered for medium_response countries
```{r}
medium_response3 <- medium_response2 %>%
  select(country, jobsat) %>%
  subset(jobsat != 'NA') %>%
  mutate(jobsat = replace(jobsat, jobsat == 'Very dissatisfied', '1-Very dissatisfied')) %>%
  mutate(jobsat = replace(jobsat, jobsat == 'Slightly dissatisfied', '2-Slightly dissatisfied')) %>%
  mutate(jobsat = replace(jobsat, jobsat == 'Neither satisfied nor dissatisfied', '3-Neutral')) %>%
  mutate(jobsat = replace(jobsat, jobsat == 'Slightly satisfied', '4-Slightly satisfied')) %>%
  mutate(jobsat = replace(jobsat, jobsat == 'Very satisfied', '5-Very satisfied'))

medium_noresponse <- dim(medium_response2[1]) - dim(medium_response3[1])
cat(medium_noresponse[1], 'respondents in medium response countries did not answer the job satisfaction question')
```

Remove entries where job satisfaction was not answered for high_response countries
```{r}
high_response3 <- high_response2 %>%
  select(country, jobsat) %>%
  subset(jobsat != 'NA') %>%
  mutate(jobsat = replace(jobsat, jobsat == 'Very dissatisfied', '1-Very dissatisfied')) %>%
  mutate(jobsat = replace(jobsat, jobsat == 'Slightly dissatisfied', '2-Slightly dissatisfied')) %>%
  mutate(jobsat = replace(jobsat, jobsat == 'Neither satisfied nor dissatisfied', '3-Neutral')) %>%
  mutate(jobsat = replace(jobsat, jobsat == 'Slightly satisfied', '4-Slightly satisfied')) %>%
  mutate(jobsat = replace(jobsat, jobsat == 'Very satisfied', '5-Very satisfied'))

high_noresponse <- dim(high_response2[1]) - dim(high_response3[1])
cat(high_noresponse[1], 'respondents in high response countries did not answer the job satisfaction question')
```

Now we'll create a barplot showing how the medium response countries rate their job satisfaction
```{r}
medium_response4 <- select(medium_response3, jobsat)

medium_jobsat <- medium_response4 %>%
  group_by(jobsat) %>%
  summarise(length(jobsat)) %>%
  rename(jobsat_count = 2)

medium_jobsat$jobsat_percentage <- round(medium_jobsat$jobsat_count/(sum(medium_jobsat$jobsat_count)),2) * 100
```

```{r}
plot_medium_jobsat <- medium_jobsat %>%
  ggplot(aes(x = jobsat, y = jobsat_percentage)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = jobsat_percentage), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 0))

plot_medium_jobsat
```

Finally, we'll create a barplot showing how the high response countries rate their job satisfaction
```{r}
high_response4 <- select(high_response3, jobsat)

high_jobsat <- high_response4 %>%
  group_by(jobsat) %>%
  summarise(length(jobsat)) %>%
  rename(jobsat_count = 2)

high_jobsat$jobsat_percentage <- round(high_jobsat$jobsat_count/(sum(high_jobsat$jobsat_count)),2) * 100
```

```{r}
plot_high_jobsat <- high_jobsat %>%
  ggplot(aes(x = jobsat, y = jobsat_percentage)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = jobsat_percentage), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 0))

plot_high_jobsat
```




## Languages

```{r import-os-csv}
survey_lang_raw <- read.csv("https://raw.githubusercontent.com/hillt5/Project-3-DATA-607/master/LanguageSkills.csv")

head(survey_lang_raw)
names(survey_lang_raw)
```
The languages dataframe contains 25 columns to rate a survey respondent's languages. I assume that the majority of people have only ranked the 3-5 languages. I'll check my assumption by counting the amount of 'NA' entries, then eliminate the unnecessary columns.


```{r lang-remove-empty-columns}
sapply(survey_lang_raw, function(x) sum(is.na(x)))
survey_lang_remove_na <- subset(survey_lang_raw, is.na(survey_lang_raw$Language1) == FALSE)
round(1-sapply(survey_lang_remove_na, function(x) sum(is.na(x))) /sum(is.na(survey_lang_remove_na$Language25)),3)
```
After removing all 'NA''s in the first language column, which coincide with non-response to this question, it appears that the last 6 columns are not contributing appreciably. The majority of responndents listed 5 languages they commonly used.

Next, I'm going to gather the Languages to see the most common overall responses were. This will sum all responses regardless of rank. I'm also going to use the facet wrap function in ggplot to 

```{r lang-plot-all}
library(tidyr)
library(dplyr)
library(ggplot2)

lang_tall <- survey_lang_remove_na %>%
  select(Language1:Language19) %>%
  gather(LangRank, Language, na.rm = TRUE)
lang_wide <- lang_tall %>%
  select(Language) %>%
  group_by(Language) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))

head(lang_wide)
nrow(lang_wide)
lang_wide[[2]][28]/lang_wide[[2]][1]

plot_lang_freq <- head(lang_wide, 10) %>%
  ggplot(aes(x = reorder(Language, Count), y = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Count), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 45))

plot_lang_freq
```

```{r lang-facet-wrap}

lang_tall_trim <- survey_lang_remove_na %>%
  select(Language1:Language10) %>%
  gather(LangRank, Language, na.rm = TRUE) 

tail(lang_tall_trim)

plot_lang_wrap_trim <- lang_tall_trim %>%
  ggplot(aes(x= Language)) +
  geom_bar() +
  facet_wrap(~LangRank)
plot_lang_wrap_trim
```
The most insteresting finding from the facet wrap is that there is one language early in the alphabet overrepresented at Language1, and two languagues in Language2 and 3 that are especially prominent. To represent this data for the presentation, I'll create plots for these three languages sepecifically.


```{r plot-first-language-preference}

lang_1_tall <- survey_lang_remove_na %>%
  select(Language1) %>%
  gather(LangRank, Language, na.rm = TRUE)
lang_1_wide <- lang_1_tall %>%
  select(Language) %>%
  group_by(Language) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))

plot_lang_1_freq <- head(lang_1_wide) %>%
  ggplot(aes(x = reorder(Language, Count), y = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Count), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 45))
plot_lang_1_freq
```

```{r plot-second-language-preference}

lang_2_tall <- survey_lang_remove_na %>%
  select(Language2) %>%
  gather(LangRank, Language, na.rm = TRUE)
lang_2_wide <- lang_2_tall %>%
  select(Language) %>%
  group_by(Language) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))

plot_lang_2_freq <- head(lang_2_wide) %>%
  ggplot(aes(x = reorder(Language, Count), y = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Count), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 45))
plot_lang_2_freq
```

```{r plot-third-language-preference}

lang_3_tall <- survey_lang_remove_na %>%
  select(Language3) %>%
  gather(LangRank, Language, na.rm = TRUE)
lang_3_wide <- lang_3_tall %>%
  select(Language) %>%
  group_by(Language) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))

plot_lang_3_freq <- head(lang_3_wide) %>%
  ggplot(aes(x = reorder(Language, Count), y = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Count), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 45))
plot_lang_3_freq
```

Looking at the top three languages, Powershell, HTML/CSS, and Javascript are the most common in the 3 language preferences. This is an unintuitive finding as overall, SQL and Python are the most common languages.  AS a final consideration, I'll take a second look the most common languages by looking only at the top language preferences and see if there are some languages with entries spread out between the top five.

```{r plot-lang-top-five}

lang_tall_5 <- survey_lang_remove_na %>%
  select(Language1:Language5) %>%
  gather(LangRank, Language, na.rm = TRUE)
lang_wide_5 <- lang_tall_5 %>%
  select(Language) %>%
  group_by(Language) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))

plot_lang_freq_top_5 <- head(lang_wide_5) %>%
  ggplot(aes(x = reorder(Language, Count), y = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Count), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 45))

plot_lang_freq_top_5
```

So while PowerShell is developers' favorite first pick, it's actually only the fourth most common pick for the top five programming languages.  HTML/CSS is the top preference in the 5 first languages, followed by Javascript, and then Python edges out Powershell as the third most common.  


## Databases

```{r import-db-csv}
survey_db_raw <- read.csv("https://raw.githubusercontent.com/hillt5/Project-3-DATA-607/master/DatabaseSkills.csv")

head(survey_db_raw)
names(survey_db_raw)
```
Databases preference contains another 13 columns, I will check the same assumption about superfluous columns.

```{r db-remove-empty-columns}
sapply(survey_db_raw, function(x) sum(is.na(x)))
survey_db_remove_na <- subset(survey_db_raw, is.na(survey_db_raw$Database1) == FALSE)

#sapply(survey_db_remove_na, function(x) sum(is.na(x))) - sapply(survey_db_raw, function(x) sum(is.na(x)))

round(1-sapply(survey_db_remove_na, function(x) sum(is.na(x))) /sum(is.na(survey_db_remove_na$Database13)),3)
```

It appears that the last row does not contain any additional information, and the last 5 contain negligible amounts.

```{r db-plot-all}
library(tidyr)
library(dplyr)
library(ggplot2)

db_tall <- survey_db_remove_na %>%
  select(Database1:Database9) %>%
  gather(DbRank, Database, na.rm = TRUE)

db_wide <- db_tall %>%
  select(Database) %>%
  group_by(Database) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))

head(db_wide)
nrow(db_wide)
db_wide[[2]][14]/db_wide[[2]][1]

plot_db_freq <- head(db_wide, 10) %>%
  ggplot(aes(x = reorder(Database, Count), y = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Count), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 45))
plot_db_freq
```

```{r db-facet-wrap}

db_tall_trim <- survey_db_remove_na %>%
  select(Database1:Database7) %>%
  gather(DbRank, Database, na.rm = TRUE) 

tail(db_tall_trim)
  
plot_db_wrap_trim <- db_tall_trim %>%
  ggplot(aes(x= Database)) +
  geom_bar() +
  facet_wrap(~DbRank)
plot_db_wrap_trim
```

Again, there appear to be three Databases clearly favored in the top three. I'll apply the same exploration and plots to Database as I did in the Language example.

```{r plot-first-db-preference}

db_1_tall <- survey_db_remove_na %>%
  select(Database1) %>%
  gather(DbRank, Database, na.rm = TRUE)
db_1_wide <- db_1_tall %>%
  select(Database) %>%
  group_by(Database) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))

plot_db_1_freq <- head(db_1_wide) %>%
  ggplot(aes(x = reorder(Database, Count), y = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Count), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 45))
plot_db_1_freq
```

```{r plot-second-db-preference}

db_2_tall <- survey_db_remove_na %>%
  select(Database2) %>%
  gather(DbRank, Database, na.rm = TRUE)
db_2_wide <- db_2_tall %>%
  select(Database) %>%
  group_by(Database) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))

plot_db_2_freq <- head(db_2_wide) %>%
  ggplot(aes(x = reorder(Database, Count), y = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Count), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 45))
plot_db_2_freq
```

```{r plot-third-db-preference}

db_3_tall <- survey_db_remove_na %>%
  select(Database3) %>%
  gather(DbRank, Database, na.rm = TRUE)
db_3_wide <- db_3_tall %>%
  select(Database) %>%
  group_by(Database) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))

plot_db_3_freq <- head(db_3_wide) %>%
  ggplot(aes(x = reorder(Database, Count), y = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Count), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 45))
plot_db_3_freq
```


```{r plot-db-top-five}

db_tall_5 <- survey_db_remove_na %>%
  select(Database1:Database5) %>%
  gather(DbRank, Database, na.rm = TRUE)
db_wide_5 <- db_tall_5 %>%
  select(Database) %>%
  group_by(Database) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))

plot_db_freq_top_5 <- head(db_wide_5) %>%
  ggplot(aes(x = reorder(Database, Count), y = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Count), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 45))

plot_db_freq_top_5
```

The number one preferred database for developers is Microsoft SQL Server, although taken as a whole MySQL and PostgreSQL are SQLite and MongoDB, and generaly there appears to be less diversity than programming languages. Only 14 databases are represented as options, as opposed to 28 Languages. The least common database is ~4% of the most commmon response, as compared to 1.9% which is the statistic for Languages.

## Platforms

```{r import-pf-csv}

survey_pf_raw <- read.csv("https://raw.githubusercontent.com/hillt5/Project-3-DATA-607/master/PlatformSkills.csv")
head(survey_pf_raw)
names(survey_pf_raw)

```

```{r pf-remove-empty-columns}

sapply(survey_pf_raw, function(x) sum(is.na(x)))

survey_pf_remove_na <- subset(survey_pf_raw, is.na(survey_pf_raw$Platform1) == FALSE)

round(1-sapply(survey_pf_remove_na, function(x) sum(is.na(x))) /sum(is.na(survey_pf_remove_na$Platform16)),3)

#sapply(survey_pf_remove_na, function(x) sum(is.na(x))) - sapply(survey_pf_raw, function(x) sum(is.na(x)))
```
Out of 16 possible platforms, most respondents picked their top three.  Platforms 12-16 represent a negligible amount of information to be gained.

```{r pf-plot-all}
library(tidyr)
library(dplyr)
library(ggplot2)

pf_tall <- survey_pf_remove_na %>%
  select(Platform1:Platform11) %>%
  gather(PfRank, Platform, na.rm = TRUE)
pf_wide <- pf_tall %>%
  select(Platform) %>%
  group_by(Platform) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))

head(pf_wide)
nrow(pf_wide)
pf_wide[[2]][17]/pf_wide[[2]][1]

plot_pf_freq <- head(pf_wide, 10) %>%
  ggplot(aes(x = reorder(Platform, Count), y = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Count), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 45))

plot_pf_freq
```

The top three database responses are Linux and Windows, with a dropoff after these two, then Docker, AWS, and MacOS.


```{r pf-facet-wrap}

pf_tall_trim <- survey_pf_remove_na %>%
  select(Platform1:Platform8) %>%
  gather(PfRank, Platform, na.rm = TRUE) 

tail(pf_tall_trim)
  
plot_pf_wrap_trim <- pf_tall_trim %>%
  ggplot(aes(x= Platform)) +
  geom_bar() +
  facet_wrap(~PfRank)
plot_pf_wrap_trim
```

The platform text wrap reveals that for the top three responses, there exist a handful of platforms that are preferred, with a significant dropoff after the third or fourth options.

```{r plot-first-pf-preference}

pf_1_tall <- survey_pf_remove_na %>%
  select(Platform1) %>%
  gather(PfRank, Platform, na.rm = TRUE)
pf_1_wide <- pf_1_tall %>%
  select(Platform) %>%
  group_by(Platform) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))

plot_pf_1_freq <- head(pf_1_wide) %>%
  ggplot(aes(x = reorder(Platform, Count), y = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Count), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 45))
plot_pf_1_freq
```

```{r plot-second-pf-preference}

pf_2_tall <- survey_pf_remove_na %>%
  select(Platform2) %>%
  gather(PfRank, Platform, na.rm = TRUE)
pf_2_wide <- pf_2_tall %>%
  select(Platform) %>%
  group_by(Platform) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))

plot_pf_2_freq <- head(pf_2_wide) %>%
  ggplot(aes(x = reorder(Platform, Count), y = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Count), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 45))
plot_pf_2_freq
```

```{r plot-third-pf-preference}

pf_3_tall <- survey_pf_remove_na %>%
  select(Platform3) %>%
  gather(PfRank, Platform, na.rm = TRUE)
pf_3_wide <- pf_3_tall %>%
  select(Platform) %>%
  group_by(Platform) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))

plot_pf_3_freq <- head(pf_3_wide) %>%
  ggplot(aes(x = reorder(Platform, Count), y = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Count), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 45))
plot_pf_3_freq
```

```{r plot-pf-top-five}

pf_tall_5 <- survey_pf_remove_na %>%
  select(Platform1:Platform5) %>%
  gather(PfRank, Platform, na.rm = TRUE)
pf_wide_5 <- pf_tall_5 %>%
  select(Platform) %>%
  group_by(Platform) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))

plot_pf_freq_top_5 <- head(pf_wide_5) %>%
  ggplot(aes(x = reorder(Platform, Count), y = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Count), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 45))

plot_pf_freq_top_5
```

Looking at the Platform reponses, it's a robust finding that Windows and Linux are the most popular, as they are represented in all of our plots, especially the more aggregated plots.  However, it's worth mentioning that Android and AWS are the top two first pick for many developers, and Docker is a preferred second pick. Again, there's more diversity in the responses for platforms as even the most uncommon answer is ~5% compared to the most common option.  

## Operating Systems

```{r}
survey_os_raw <- read.csv("https://raw.githubusercontent.com/hillt5/Project-3-DATA-607/master/OpSysSkills.csv")

head(survey_os_raw)
names(survey_os_raw)
```

This portion looks fairly clean, and is already in a tall format

```{r}
sapply(survey_os_raw, function(x) sum(is.na(x)))
survey_os_remove_na <- subset(survey_os_raw, is.na(survey_os_raw$OpSys) == FALSE)

sapply(survey_os_remove_na, function(x) sum(is.na(x)))
```

```{r}
os_wide <- survey_os_remove_na %>%
  select(OpSys) %>%
  group_by(OpSys) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))


plot_os_freq <- os_wide %>%
  ggplot(aes(x = reorder(OpSys, Count), y = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Count), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 45))
plot_os_freq
```

It appears Windows is by far the most popular operating system, followed by Linux. MacOS comes in third, which may be surprising at first, given the popularity of Macs in the uNited States, but this is a global survey, which may play a factor in these rankings.

```{r}
demo_raw <- read.csv('https://raw.githubusercontent.com/hillt5/Project-3-DATA-607/master/Demographics.csv')

sapply(demo_raw, function(x) sum(is.na(x)))
demo_remove_na <- subset(demo_raw, is.na(demo_raw$Respondent) == FALSE)

nrow(survey_os_raw)
nrow(demo_raw)
```

I will investigate if the countries have an effect on the OS numbers

```{r}
os_countries <- left_join(survey_os_raw, demo_raw, by = "Respondent")
keep <- c("Respondent", "OpSys", "Country")
os_countries <- os_countries[keep]
head(os_countries)
```
```{r}
# drop NA
sapply(os_countries, function(x) sum(is.na(x)))
os_countries <- subset(os_countries, is.na(os_countries$OpSys) == FALSE && is.na(os_countries$Country) == FALSE)

# consolidate to US and International

os_intl <- os_countries %>%
  select(Country, OpSys) %>%
  subset(OpSys != 'NA') %>%
  mutate(Country = as.character(Country)) %>%
  mutate(Country = replace(Country, Country != 'United States', 'International'))
  

#for(i in 1:nrow(os_countries)){
  #if(os_countries$Country[i] != "United States"){
   # os_countries$Country[i] <- "International"
  #}}


head(os_intl, 15)

os_intl_wide <- os_intl %>%
  select(OpSys, Country) %>%
  group_by(Country, OpSys) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count, Country))
os_intl_wide

plot_os_intl <- os_intl_wide %>%
  ggplot(aes(x = reorder(OpSys, Count), y = Count)) + 
  geom_bar(stat = "identity") + 
  facet_wrap(~Country)+
  geom_text(aes(label = Count), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 45))
plot_os_intl
```


## WebFrames

```{r}
survey_wf_raw <- read.csv("https://raw.githubusercontent.com/hillt5/Project-3-DATA-607/master/WebFrameSkills.csv")

head(survey_wf_raw)
names(survey_wf_raw)
```
There seems to be 12 webframe values. We will see how many contain mostly null values.

```{r}
sapply(survey_wf_raw, function(x) sum(is.na(x)))
survey_wf_remove_na <- subset(survey_wf_raw, is.na(survey_wf_raw$WebFrame1) == FALSE)

round(1-sapply(survey_wf_remove_na, function(x) sum(is.na(x))) /sum(is.na(survey_wf_remove_na$WebFrame12)),3)
```
It appears that the majority of respondents use no more than 2 web frameworks. Given this, it seems reasonable to limit this to the top 4.

```{r}
library(tidyr)
library(dplyr)
library(ggplot2)

wf_tall <- survey_wf_remove_na %>%
  select(WebFrame1:WebFrame4) %>%
  gather(WfRank, WebFrame, na.rm = TRUE)

wf_wide <- wf_tall %>%
  select(WebFrame) %>%
  group_by(WebFrame) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))

head(wf_wide)
nrow(wf_wide)
wf_wide[[2]][12]/wf_wide[[2]][1]

plot_wf_freq <- head(wf_wide, 10) %>%
  ggplot(aes(x = reorder(WebFrame, Count), y = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Count), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 45))
plot_wf_freq
```

jQuery is the runaway favorite web framework, followed by Angular, now lets looks at the top two.

```{r}
wf_tall_trim <- survey_wf_remove_na %>%
  select(WebFrame1:WebFrame2) %>%
  gather(WfRank, WebFrame, na.rm = TRUE)

tail(wf_tall_trim)
  
plot_wf_wrap_trim <- wf_tall_trim %>%
  ggplot(aes(x= WebFrame)) +
  geom_bar() +
  facet_wrap(~WfRank)+
  theme(axis.text.x = element_text(angle = 90))
plot_wf_wrap_trim
```

Interestingly, Angular appears to be the most popular first, and jQuery picks up many tallies as the second choice of many developers. JavaScript frameworks are heavily represented in the top choices. It is also interesting that Angular does not appear as a second choice.

## Development Environment SKills

```{r}
survey_de_raw <- read.csv("https://raw.githubusercontent.com/hillt5/Project-3-DATA-607/master/DevEnvironSkills.csv")

head(survey_de_raw)
names(survey_de_raw)
```

There seem to be 22 DevEnviron columns. We will see how how contain significant information.

```{r}
sapply(survey_de_raw, function(x) sum(is.na(x)))
survey_de_remove_na <- subset(survey_de_raw, is.na(survey_de_raw$DevEnviron1) == FALSE)

round(1-sapply(survey_de_remove_na, function(x) sum(is.na(x))) /sum(is.na(survey_de_remove_na$DevEnviron22)),3)
```

Since nearly 90% of respondents listed 5 or fewer environments, we will trim this to the top 5.

```{r}
de_tall <- survey_de_remove_na %>%
  select(DevEnviron1:DevEnviron5) %>%
  gather(DeRank, DevEnviron, na.rm = TRUE)

de_wide <- de_tall %>%
  select(DevEnviron) %>%
  group_by(DevEnviron) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))

plot_de_freq <- head(de_wide, 12) %>%
  ggplot(aes(x = reorder(DevEnviron, Count), y = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Count), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 90))
plot_de_freq
```

Visual Studio Code has the most appearances in the top five. Jupyter is in thrid and RStudio is in 12th. Let look at the top preferences.

```{r}
de_1_tall <- survey_de_remove_na %>%
  select(DevEnviron1) %>%
  gather(DeRank, DevEnviron, na.rm = TRUE)

de_1_wide <- de_1_tall %>%
  select(DevEnviron) %>%
  group_by(DevEnviron) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))

plot_de_1_freq <- head(de_1_wide, 12) %>%
  ggplot(aes(x = reorder(DevEnviron, Count), y = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Count), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 90))
plot_de_1_freq
```
Interesetingly, Visual Studio Code falls to 11th in first choices,depite having the most occurences in the top five, while Android Studio or Jupyter are the top picks for most people. Let's see second choices.

```{r}
de_2_tall <- survey_de_remove_na %>%
  select(DevEnviron2) %>%
  gather(DeRank, DevEnviron, na.rm = TRUE)

de_2_wide <- de_2_tall %>%
  select(DevEnviron) %>%
  group_by(DevEnviron) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))

plot_de_2_freq <- head(de_2_wide, 12) %>%
  ggplot(aes(x = reorder(DevEnviron, Count), y = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Count), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 90))
plot_de_2_freq
```
We see that Notepad++ is a popular second choice, and Jupyter reamins a strong second as a second choice. RStudio was the 10th most popular second choice.

```{r}
de_3_tall <- survey_de_remove_na %>%
  select(DevEnviron3) %>%
  gather(DeRank, DevEnviron, na.rm = TRUE)

de_3_wide <- de_3_tall %>%
  select(DevEnviron) %>%
  group_by(DevEnviron) %>%
  summarize(Count = n()) %>%
  arrange(desc(Count))

plot_de_3_freq <- head(de_3_wide, 12) %>%
  ggplot(aes(x = reorder(DevEnviron, Count), y = Count)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Count), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 90))
plot_de_3_freq
```