---
title: "Project 3 - Demographics"
author: "Ken Popkin"
date: "3/21/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```


```{r}
library(RMySQL)
library(dplyr)
library(ggplot2)
```


### Read from AWS MySQL database
```{r}
db = dbConnect(MySQL(), user='root', password='masterfightingduckhorse', dbname='project3', host='mydbinstance.c1u7dqmta5za.us-east-1.rds.amazonaws.com', port=3306)
```

```{r}
rs = dbSendQuery(db, "select * from demographics")
dfdemographics = fetch(rs, n=-1)
```

```{r}
dfdemographics <- select (dfdemographics,-c(dummy))
dim(dfdemographics)
```
 
We'll now take a look at job satisfaction for countries with a medium (100 - 999 respondents) and high (> 1000 respondents)
```{r}
dfcountry <- select(dfdemographics, country)

response <- dfcountry %>%
  group_by(country) %>%
  summarise(length(country)) %>%
  rename(respondents = 2)
  
medium_response <- subset(response, respondents > 99 & respondents < 501)
high_response <- subset(response, respondents > 999)

medium_response_vector <- pull(medium_response,country)
high_response_vector <- pull(high_response,country)

cat('Medium response countries are:', medium_response_vector, '\n\n')
cat('High response countries are:', high_response_vector)
```

```{r}
#Dataframe of medium response countries job satisfaction ratings
medium_response2 <- filter(dfdemographics, country %in% medium_response_vector)
dim(medium_response2)

#Dataframe of high response countries job satisfaction ratings
high_response2 <- filter(dfdemographics, country %in% high_response_vector)
dim(high_response2)
```

Remove entries where job satisfaction was not answered for medium_response countries
```{r}
medium_response3 <- medium_response2 %>%
  select(country, jobsat) %>%
  subset(jobsat != 'NA') %>%
  mutate(jobsat = replace(jobsat, jobsat == 'Neither satisfied nor dissatisfied', 'Neutral'))

medium_noresponse <- dim(medium_response2[1]) - dim(medium_response3[1])
cat(medium_noresponse[1], 'respondents in medium response countries did not answer the job satisfaction question')
```

Remove entries where job satisfaction was not answered for high_response countries
```{r}
high_response3 <- high_response2 %>%
  select(country, jobsat) %>%
  subset(jobsat != 'NA') %>%
  mutate(jobsat = replace(jobsat, jobsat == 'Neither satisfied nor dissatisfied', 'Neutral'))

high_noresponse <- dim(high_response2[1]) - dim(high_response3[1])
cat(high_noresponse[1], 'respondents in high response countries did not answer the job satisfaction question')
```

Now we'll create a barplot showing how the medium response countries rate their job satisfaction
```{r}
medium_response4 <- select(medium_response3, jobsat)

medium_jobsat <- medium_response4 %>%
  group_by(jobsat) %>%
  summarise(length(jobsat)) %>%
  rename(jobsat_count = 2)
```

```{r}
plot_medium_jobsat <- medium_jobsat %>%
  ggplot(aes(x = jobsat, y = jobsat_count)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = jobsat_count), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 45))

plot_medium_jobsat
```

Finally, we'll create a barplot showing how the high response countries rate their job satisfaction
```{r}
high_response4 <- select(high_response3, jobsat)

high_jobsat <- high_response4 %>%
  group_by(jobsat) %>%
  summarise(length(jobsat)) %>%
  rename(jobsat_count = 2)
```

```{r}
plot_high_jobsat <- high_jobsat %>%
  ggplot(aes(x = jobsat, y = jobsat_count)) + 
  geom_bar(stat = "identity") + 
  geom_text(aes(label = jobsat_count), vjust = -0.3) +
  theme(axis.text.x = element_text(angle = 45))

plot_high_jobsat
```